Learning Discrete-valued Bayesian Networks from Mixed Data

### 2015-11-17 BN_discretizer_paper.tex
looking at variable definitions

\newcommand{\todo}[1]{\textcolor{magenta}{#1}}
\newcommand{\tim}[1]{\textit{\textcolor{blue}{#1}}}

@@

\section{Introduction}
\label{intro}
Bayesian networks \citep{Pearl_1988, PGM_2009} are an increasingly popular method for modeling uncertainty and causality in science and engineering. They provide an efficient factorization of the joint probability distribution over a set of random variables. Bayesian networks first emerged from artificial intelligence research and have been applied to a wide variety of problems, ranging from decision-making systems \citep{DMU_2015} to medical diagnoses \citep{Lustgarten_2011}. In most cases, we assume that all random variables in Bayesian networks are discrete, since many Bayesian network learning algorithms are unable to deal with continuous variables efficiently. However, many applications
require the use of continuous variables, such as position and velocity in dynamic systems.

There are two common approaches to extending Bayesian networks to continuous variables.
The first is to model the conditional probability density of each continuous variable using specific families of parametric distributions, and then to redesign Bayesian network learning algorithms based on their parameters.
One successful example is belief propagation in Gaussian graphical models \citep{Weiss_2011}.
For other parametric distributions \citep{Ihler_2009} and non-linear functions~\todo{[cite]}, these redesigned algorithms are computationally expensive and do not perform well.

@@

\subsection{MDL Principle Discretization}
The MDL principle was first proposed by \cite{MDL_1978}, and states that the best model for a dataset is one which minimizes the amount of information needed to describe it \citep{Grunwald_2009}.
MDL methods trade off goodness-of-fit against model complexity to reduce generalization error.
In the context of Bayesian networks, \cite{Friedman_1996} applied the MDL principle to determine the optimal number of discretization intervals for continuous variables and optimal the positions of their discretization boundaries.
Their approach selects a discretization policy that minimizes the sum of the description lengths of the discretized Bayesian network and the information necessary for recovering the continuous values from the discretized data.
In the single continuous variable case, the objective function is

@@

\noindent
% $I(\boldsymbol{A},\boldsymbol{B}) =  \sum_{\boldsymbol{a},\boldsymbol{b}} \hat{P}_D (\boldsymbol{a},\boldsymbol{b}) \log {{\hat{P}_D (\boldsymbol{a},\boldsymbol{b})}\over{\hat{P}_D (\boldsymbol{a}),\hat{P}_D (\boldsymbol{b})}}$
where $I(\boldsymbol{A},\boldsymbol{B})$ is the mutual information between the discrete variables $\boldsymbol{A}$ and $\boldsymbol{B}$, ${H(p) = -p \log(p) - (1-p) \log(1-p)}$, $|\boldsymbol{X}|$ is the cardinality of discrete variable $\boldsymbol{X}$, and $X^*_i$ is the discretized version of the $i$th sample of continuous variable $X$.
The optimal discretization policy can be found with dynamical programming.
For Bayesian networks with multiple continuous variables, the algorithm is applied to one variable at a time iteratively over all continuous variables.
While iterating, only one variable is treated as continuous and other continuous variables are discretized based on an initial discretization policy or the discretization result from a previous iteration.

@@

% The discretization technique proposed by \citet{Friedman_1996} is the most well-known of these methods. It is based on the minimum description length principle (MDL): the optimal discretization policy minimizes the description length of the Bayesian network and discretization policy. The MDL algorithm has a $\bigo\paren{N^3 + \paren{n_c {v_\text{max}}^{(n_c^p)_\text{max}} + {v_\text{max}}^{n_p}} \cdot N^2}$ runtime on Bayesian networks with a single continuous variable and multiple discrete variables, where $N$ is the number of data samples from the joint distribution, $n_p$ and $n_c$ are the numbers of parent and children variables for the continuous variable, $v_\text{max}$ is the largest cardinality over all variables in the Markov blanket, and  ${(n_c^p)_\text{max}}$ is the largest number of parent variables of the continuous variable's children. The MDL method iterates over all continuous variables until convergence in the case of multiple continuous variables, converging in a few cycles according to our tests of real-world data. Unfortunately, MDL tends to produce discrete variables with low cardinality~\todo{[cite]}.

@@

The MODL method uses dynamic programming to find the optimal discretization policy, and has a $\bigo\paren{N^3 + |\text{Class}| \cdot N^2}$ runtime, where $N$ is the of number of data instances.
\cite{Lustgarten_2011} suggest several formulations for the prior term $P(\text{Model})$.
Under a uniform prior probability over discretizations the runtime can be reduced to $\bigo\paren{|\text{Class}| \cdot N^2}$ without sacrificing optimality.

@@

A Bayesian network $B$ is defined by a pair $(G,\Theta)$, where $G = (X,E)$ is a directed acyclic graph whose nodes correspond to a set of random variables $X = \brock{X_1, X_2, \cdots, X_n}$, and whose edges $E$ represent probabilistic dependencies among nodes. The graph structure $G$ encodes the Markov property: each node $X_i$ is independent of its non-descendants given its parents $\Pi_{X_i}$ in $G$. The conditional relations are parameterized by the elements of $\Theta$, which take the form $\theta_{x_i \mid \Pi_{x_i}} = P(x_i \mid \Pi_{x_i})$ for each possible value $x_i$ of $X_i$, and $\Pi_{x_i}$ of $\Pi_{X_i}$. The joint distribution over $X$ factors according to the Markov property:
\begin{equation}
P_B (X_1 , \cdots, X_n) = \prod_{i=1}^{n} P_B (X_i \mid \Pi_{X_i}) = \prod_{i=1}^{n} \theta_{x_i \mid \Pi_{x_i}}\text{.}
\end{equation}

Forward sampling is used to draw samples from the joint distribution represented by a Bayesian network.
Variables are sampled in topological order such that parental instantiations are always known~\citep[see][chap.~22]{algo_2009}.
Furthermore, the parent $\rightarrow$ child sampling order can be reversed if the marginal probability of the child variable is known.
By Bayes' rule,

\begin{equation}
P(X_i \mid \Pi_{X_i}) \cdot \prod_{j = 1}^{ Pa( X_i)} P( \{ \Pi_{X_i} \}_j) = P(X_i) \cdot P(\Pi_{X_i} \mid X_i),
\end{equation}

\noindent
where $Pa(X_i)$ is the number of parents of $X_i$ and $\brock{\Pi_{X_i}}_j$ is the $j$th parent of $X_i$.
% One can first sample $X_i$ by $P(X_i)$, then sample all the parent variables simultaneously by $P(\Pi_{X_i} \mid X_i)$. That is to say, $X_i$ becomes the starting point of sampling.

@@

A discretization policy $M_C = \anglebrackets{t_1 < t_2 < \ldots < t_{k-1}}$ for a continuous variable $C$ is a mapping from $\mathbb{R}$ to $\brock{1,2,3,\ldots,k}$ such that
\begin{equation}
  M_C (x)=\begin{cases}
    1, & \text{if $x<t_1$}.\\
    i, & \text{if $t_{i-1} \leq x < t_i$}\text{.}\\
    k, & \text{if $t_{k-1} \leq x$}
  \end{cases}
\end{equation}

\noindent
The policy discretizes the continuous variable $C$ into $k$ intervals.
The discretization edges $t_\cdot$ are restricted to the midpoints between two sorted instances of $C$.
An integer representation of $M_C$ can be obtained by sorting the samples of $C$ in an increasing order, $\curly{c_1,c_2,\ldots,c_N}$, and if $t_i = (c_{s_{i}} + c_{s_{i}+1})/2$ for $i=1,2,\ldots,k$, then $M_C = \anglebrackets{t_1,t_2, \ldots, t_{k-1}} \equiv \brock{n_1,n_2,..n_k}$, where $n_1 = s_1$ and $n_i = s_{i} - s_{i-1}$.

@@

\section{Discretization of a Single Continuous Variable}
\label{sec:single_var}

@@

Let $D$ be a dataset of $N$ samples from which the discretization will be learned, sorted in ascending order according to $X$.
Let $D_{\boldsymbol{Y}}$ will refer to the data instances associated with variables $\boldsymbol{Y}$, thus $D_X = \{ x_1,x_2,x_3,\ldots,x_N \}$.
In order to facilitate understanding, we temporarily assume all values in $D_X$ are unique, that is, $x_1 < x_2 < \ldots < x_N$.
It follows that a discretization policy $M$ on $X$ can be written as $M_X = \brock{n_1,n_2,\ldots,n_k}$, where $k$, $n_1$, $n_2$, $\ldots,n_k$ are positive integers satisfying $N = \sum_{i=1}^k n_i$.
An optimal discretization policy maximizes $P(M)\cdot P(D\mid M)$ for some prior $P(M)$ and likelihood $P(D\mid M)$.

@@

The following four principles for the optimal discretization policy, modivated by MODL, enable the formulation of $P(M)$ and $P(D \mid M)$:
\begin{enumerate}
\item A discretization edge located at $(x_i + x_{i+1})/2$ has probability
\begin{equation}
1 - {\exp(- L \cdot {{x_{i+1} - x_i}\over{x_N - x_1}})},
\end{equation}
where $L$ is the largest cardinality number of discrete variables in $X$'s Markov blanket. \tim{Where does this come from? Is there an intuitive explanation?}
\item For a given interval of $M$, every distribution of $\Pi_X$ is equiprobable.
\item For each pair of $(C_i,\boldsymbol{S_i})$ and a given interval of $M$, every distribution of $C_i$ with a value of $\boldsymbol{S_i}$ is equiprobable.
\item The distributions of each $C_i$ and $\Pi_X$ in each interval of $M$ with each value of $\boldsymbol{S_i}$ are independent from each other.
\end{enumerate}

From the first principle one obtains
\begin{equation}
P(M) = \prod_{i=1}^k \curly{ \brock{1 - {\exp(- L \cdot {{x_{s_i+1} - x_{s_i}}\over{x_N - x_1}})}} \cdot \exp\paren{L \cdot {{x_{s_{i}} - x_{s_{i-1} + 1}}\over{x_N - x_1}}}},
\end{equation}

@@

\begin{equation}
P\paren{D \mid M} \propto P\paren{D_{\Pi_x} \mid M} \cdot \prod_{i = 1}^{n_c} P\paren{D_{C_i} \mid M, D_{\boldsymbol{S_i}}}\text{.}
\end{equation}

@@

\noindent
where $n^{(j)}_{i,m,l}$ is the number of instances in $i$th interval with $m$th value of $C_j$ and $l$th value of $\boldsymbol{S_j}$, $n_{i,l} = \sum_{m=1}^{J_j} n^{(j)}_{i,m,l}$, and $n_i = \sum_{l=1}^{L_j} n_{i,l}$.
The two factors on the right hand side come from the third prior: all distribution of values of $C_j$ in a given interval and with a given value of $S_j$ are equiprobable. According to the forth prior, these distributions are independent from each other, and one thus takes their product. If $S_j = \emptyset$, then Equation \ref{eq:likelihood_two} is equivalent to

@@

The objective function can be formulated given equations \ref{eq:p_M}, \ref{eq:p_D_given_M},\ref{eq:likelihood_one}, and \ref{eq:likelihood_three}.
The log-inverse of $P(M) \cdot P(D|M)$ is minimized for computational convenience:

@@

Now we are able to solve the optimization problem over Equation \ref{eq:opt_prob}.
The dynamic programming procedure is shown in Algorithm \ref{alg:disc_one}.
It takes two inputs: $D$, the joint samples over all variables sorted in ascending order according to $D_X$, and $G$, the network structure.
The runtime of Algorithm \ref{alg:disc_one} is also $O(n_c  {v'_\text{max}} \cdot N^2 + {v'_\text{max}}^{n_p} \cdot N^2)$, because the runtime of the dynamic programming procedure is less than the runtime for computing $h(u,v)$.
Algorithm \ref{alg:disc_one} is guaranteed to be optimal.
For faster methods with suboptimal results please refer to \citep{Boulle_2006}.

@@

Algorithm \ref{alg:disc_one} has an exponential $\bigo\paren{\paren{v'_\text{max}}^{n_p} \cdot N^2}$ runtime which severely limits discretization in networks with a large number of discretization intervals or parent variables.
This section introduces an approximation to the objective function that significantly reduces the runtime and still preserves the quality of discretization.
The approximation replaces the dominator of the last factor in Equation \ref{eq:likelihood_one} with

@@

The approximation reduces the runtime of Algorithm \ref{alg:disc_one} to a polynomial, $\bigo\paren{ {v'_\text{max}} \cdot (n_c + n_p) \cdot N^2}$.
It will be shown in Section \ref{sec:experiments} that the approximation is more sensitive to the distribution over other variables and is biased towards slightly higher interval counts.
\tim{We need to have a more specific thing to say along the lines of When does it fail?}

With the approximated objective function, Equation \ref{eq:opt_prob_approx}, provides intuition into how child and parent variables contribute to the objective function.
For example, in the left graph of Figure \ref{fig:example_networks}, the corresponding square brackets in the approximated objective function are

@@

\begin{algorithm}
  \label{alg:structure_learn}
  \caption{ Learning a discrete-valued Bayesian network}
  \begin{algorithmic}[5]
  \Function{Learn\_DBN}{$D''$, $C$, $order$, $u_{parent}$, $C$, $u_{cycle}$}
    \State $n \leftarrow$ the number of variables in the Bayesian network
    \State $k \leftarrow median\{ \|v\|, v\notin C\}$
    \State $M[i] \leftarrow$ the discretization policy for the $i$th variable
    \State $D^*_i \leftarrow M[i] (D_i)$, list of samples $D_i$ discretized according to $M[i]$
    \State $D^* \leftarrow $ the discretized data for all variables, where $D^*_i \leftarrow D''_i$ if $i \notin C$
    \State $G \leftarrow$ the initial edgeless graph structure
    \For {$i = 1$ to $n$}
      \If {$i \in C$}
        \State $M[i] \leftarrow$  equal-width discretization with $k$ intervals
        \State $D^*_i \leftarrow  M[i] (D''_i)$
      \EndIf
    \EndFor
    \For {$i =1$ to $n$}
      \State $P_{old} \leftarrow f(X^*_i,\Pi_{X^*_i})$: Equation \ref{eq:f_x_p}
      \State OKToProceed $\leftarrow$ \textbf{true}
      \While {OKToProceed \textbf{and} $\|\Pi_{X^*_i}\| < u_{parent}$}
        \State $Y \leftarrow$ an element from the set $order[1:i] \backslash \Pi_X$
        \State $P_{new} \leftarrow f(X^*_i,\Pi_{X^*_i} \cup Y)$
        \If {$P_{new} > P_{old}$}
          \State $P_{old} \leftarrow P_{new}$
          \State $\Pi_{X_i} \leftarrow \Pi_{X_i} \cup Y $
          \State $M \leftarrow$ {DISCRETIZE}({$D''$, $G$, $C$, $u_{cycle}$}): Algorithm \ref{alg:disc_two}
          \State $D* = M(D'')$
        \Else
          \State OKToProceed $\leftarrow$ \textbf{false}
        \EndIf
      \EndWhile
    \EndFor
    \State \Return $G$, $M$
  \EndFunction
  \end{algorithmic}
\end{algorithm}

>>>>

\newcommand{\todo}[1]{\textcolor{magenta}{#1}}
\newcommand{\tim}[1]{\textit{\textcolor{blue}{#1}}}
\newcommand{\mykel}[1]{\textit{\textcolor{cyan}{#1}}}

@@

\section{Introduction}
\label{intro}
Bayesian networks \citep{Pearl_1988, PGM_2009} are an increasingly popular method for modeling uncertainty and causality in science and engineering. They provide an efficient factorization of the joint probability distribution over a set of random variables. Bayesian networks first emerged from artificial intelligence research and have been applied to a wide variety of problems, ranging from decision-making systems \citep{DMU_2015}\mykel{better to use a primary source} to medical diagnoses \citep{Lustgarten_2011}. In most cases, we assume that all random variables in Bayesian networks are discrete, since many Bayesian network learning algorithms are unable to deal with continuous variables efficiently. However, many applications
require the use of continuous variables, such as position and velocity in dynamic systems \mykel{cite Kochenderfer JGCD article on encounter modeling}.

There are two common approaches to extending Bayesian networks to continuous variables.
The first is to model the conditional probability density of each continuous variable using specific families of parametric distributions, and then to redesign Bayesian network learning algorithms based on their parameters.
One successful example is belief propagation in Gaussian graphical models \citep{Weiss_2011}.
For other parametric distributions \citep{Ihler_2009} and non-linear functions~\todo{[cite]}, these redesigned algorithms are computationally expensive and do not perform well.
\mykel{$\bigstar$ We need to discuss this; talk about algorithms for inference and structure learning, expectation propagation, Gibbs, BVGS}

@@

\subsection{MDL Discretization}
The MDL principle was first proposed by \cite{MDL_1978} and states that the best model for a dataset is one that minimizes the amount of information needed to describe it \citep{Grunwald_2009}.
MDL methods trade off goodness-of-fit against model complexity to reduce generalization error.
In the context of Bayesian networks, \cite{Friedman_1996} applied the MDL principle to determine the optimal number of discretization intervals for continuous variables and the optimal positions of their discretization boundaries.
Their approach selects a discretization policy that minimizes the sum of the description lengths of the discretized Bayesian network and the information necessary for recovering the continuous values from the discretized data.
For a Bayesian network with a single continuous variable $X$, the objective function is

@@

\mykel{$\bigstar\Pi_{X_i}$ is not defined until later}
\tim{$\Pi_{X_i}$ may be confusing with our use of $\prod_{i=1}^n$ later on.}

\noindent
% $I(\boldsymbol{A},\boldsymbol{B}) =  \sum_{\boldsymbol{a},\boldsymbol{b}} \hat{P}_D (\boldsymbol{a},\boldsymbol{b}) \log {{\hat{P}_D (\boldsymbol{a},\boldsymbol{b})}\over{\hat{P}_D (\boldsymbol{a}),\hat{P}_D (\boldsymbol{b})}}$
where $I(A,B)$ is the mutual information between the discrete variables $A$ and $B$, ${H(p) = -p \log(p) - (1-p) \log(1-p)}$ \tim{does $H(p)$ have a name? Entropy?}, $|A|$ is the cardinality of a discrete variable $A$, and $X^*_i$ is the discretized version of the $i$th sample of $X$.
\tim{What are $N$, and $N_i$?}
The optimal discretization policy can be found with dynamical programming.
For Bayesian networks with multiple continuous variables, the algorithm is applied to one variable at a time iteratively over all continuous variables.
While iterating, only one variable is treated as continuous and other continuous variables are discretized based on an initial discretization policy or the discretization result from a previous iteration.

@@

% The discretization technique proposed by \citet{Friedman_1996} is the most well-known of these methods. It is based on the minimum description length principle (MDL): the optimal discretization policy minimizes the description length of the Bayesian network and discretization policy. The MDL algorithm has an $\bigo\paren{N^3 + \paren{n_c {v_\text{max}}^{(n_c^p)_\text{max}} + {v_\text{max}}^{n_p}} \cdot N^2}$ runtime on Bayesian networks with a single continuous variable and multiple discrete variables, where $N$ is the number of data samples from the joint distribution, $n_p$ and $n_c$ are the numbers of parent and children variables for the continuous variable, $v_\text{max}$ is the largest cardinality over all variables in the Markov blanket, and  ${(n_c^p)_\text{max}}$ is the largest number of parent variables of the continuous variable's children. The MDL method iterates over all continuous variables until convergence in the case of multiple continuous variables, converging in a few cycles according to our tests of real-world data. Unfortunately, MDL tends to produce discrete variables with low cardinality~\todo{[cite]}\mykel{don't like low-cardinality. Maybe 'intervals' or something instead}.

@@

The MODL method uses dynamic programming to find the optimal discretization policy, and has an $\bigo\paren{N^3 + |\text{Class}| \cdot N^2}$ runtime, where $N$ is the of number of data instances.
\cite{Lustgarten_2011} suggest several formulations for the prior term $P(\text{Model})$.
Under a uniform prior probability over discretizations the runtime can be reduced to $\bigo\paren{|\text{Class}| \cdot N^2}$ without sacrificing optimality.

@@

\mykel{This all needs to be explained better. Too many variables are being defined at once.}
A Bayesian network $B$ is defined by a pair $(G,\Theta)$, where $G = (X,E)$ is a directed acyclic graph whose nodes correspond to a set of random variables $X = \brock{X_1, X_2, \ldots, X_n}$, and whose edges $E$ represent probabilistic dependencies among nodes.
The graph structure $G$ encodes the Markov property: each node $X_i$ is independent of its non-descendants given its parents $\Pi_{X_i}$ in $G$.
The conditional relations are parameterized by the elements of $\Theta$, which take the form $\theta_{x_i \mid \Pi_{x_i}} = P(x_i \mid \Pi_{x_i})$ for each possible value $x_i$ of $X_i$, and $\Pi_{x_i}$ of $\Pi_{X_i}$. The joint distribution over $X$ factors according to the Markov property:

\begin{equation}
P_B (X_1 , \ldots, X_n) = \prod_{i=1}^{n} P_B (X_i \mid \Pi_{X_i}) = \prod_{i=1}^{n} \theta_{x_i \mid \Pi_{x_i}}\text{.}
\end{equation}

Forward sampling is used to draw samples from the joint distribution represented by a Bayesian network.
Variables are sampled in topological order such that parental instantiations are always known~\citep[see][chap.~22]{algo_2009}.
\mykel{I think this ref is just for the topological sort; for sampling, cite DMU or PGM}
Furthermore, the parent $\rightarrow$ child sampling order can be reversed if the marginal probability of the child variable is known.
By Bayes' rule,

\begin{equation}
P(X_i \mid \Pi_{X_i}) \cdot \prod_{j = 1}^{ Pa( X_i)} P( \brock{\Pi_{X_i} }_j) = P(X_i) \cdot P(\Pi_{X_i} \mid X_i),
\end{equation}

\noindent
where $Pa(X_i)$ is the number of parents of $X_i$ and $\brock{\Pi_{X_i}}_j$ is the $j$th parent of $X_i$.
\mykel{this may be messed up. Maybe use DMU notation?}
% One can first sample $X_i$ by $P(X_i)$, then sample all the parent variables simultaneously by $P(\Pi_{X_i} \mid X_i)$. That is to say, $X_i$ becomes the starting point of sampling.

@@

\tim{We need to adopt the same notation everywhere; $C$ should be a child, not a continuous variable}.
A discretization policy $M_C = \anglebrackets{t_1 < t_2 < \ldots < t_{k-1}}$ for a continuous variable $C$ is a mapping from $\mathbb{R}$ to $\brock{1,2,3,\ldots,k}$ such that
\begin{equation}
  M_C (x)=\begin{cases}
    1, & \text{if $x<t_1$}.\\
    i, & \text{if $t_{i-1} \leq x < t_i$}\text{.}\\
    k, & \text{if $t_{k-1} \leq x$}
  \end{cases}
\end{equation}

\noindent
The policy discretizes the continuous variable $C$ into $k$ intervals.
The discretization edges $t_{1:k-1}$ are restricted to the midpoints between two sorted instances of $C$.
An integer representation of $M_C$ can be obtained by sorting the samples of $C$ in an increasing order, $\curly{c_1,c_2,\ldots,c_N}$, and if $t_i = (c_{s_{i}} + c_{s_{i}+1})/2$ for $i=1,2,\ldots,k$, then $M_C = \anglebrackets{t_1,t_2, \ldots, t_{k-1}} \equiv \brock{n_1,n_2,..n_k}$, where $n_1 = s_1$ and $n_i = s_{i} - s_{i-1}$.
\mykel{Where did $s_i$ come from?}

@@

\section{Single Variable Discretization}
\label{sec:single_var}

@@

Let $D$ be a dataset of $N$ samples from which the discretization will be learned, sorted in ascending order according to $X$.
Let $D_{\boldsymbol{Y}}$ will refer to the data instances associated with variables $\boldsymbol{Y}$, thus $D_X = \{ x_1,x_2,x_3,\ldots,x_N \}$.
In order to facilitate understanding, we temporarily assume all values in $D_X$ are unique, that is, $x_1 < x_2 < \ldots < x_N$.
It follows that a discretization policy $M$ on $X$ can be written as $M_X = \brock{n_1,n_2,\ldots,n_k}$, where $k$, $n_1$, $n_2$, $\ldots,n_k$ are positive integers satisfying $N = \sum_{i=1}^k n_i$.
\mykel{$\bigstar$ This conflicts with Section 3.2, right?}
An optimal discretization policy maximizes $P(M)\cdot P(D\mid M)$ for some prior $P(M)$ and likelihood $P(D\mid M)$.

@@

The following four principles for the optimal discretization policy, motivated by MODL, enable the formulation of $P(M)$ and $P(D \mid M)$:
\begin{enumerate}
\item A discretization edge located at $(x_i + x_{i+1})/2$ has probability
\begin{equation}
1 - {\exp(- L \cdot {{x_{i+1} - x_i}\over{x_N - x_1}})},
\end{equation}
where $L$ is the largest cardinality number of discrete variables in $X$'s Markov blanket. \tim{Where does this come from? Is there an intuitive explanation?}\mykel{$\bigstar$}
\item For a given interval of $M$, every distribution of $\Pi_X$ is equiprobable.
\item For each pair of $(C_i,\boldsymbol{S_i})$ and a given interval of $M$, every distribution of $C_i$ with a value of $\boldsymbol{S_i}$ is equiprobable.\mykel{$\bigstar$}
\item The distributions of each $C_i$ and $\Pi_X$ in each interval of $M$ with each value of $\boldsymbol{S_i}$ are independent from each other.\mykel{$\bigstar$}
\end{enumerate}

From the first principle one obtains
\begin{equation}
\label{eq:p_M}
P(M) = \prod_{i=1}^k \curly{ \brock{1 - {\exp(- L \cdot {{x_{s_i+1} - x_{s_i}}\over{x_N - x_1}})}} \cdot \exp\paren{L \cdot {{x_{s_{i}} - x_{s_{i-1} + 1}}\over{x_N - x_1}}}},
\end{equation}

@@

\begin{equation}
\label{eq:p_D_given_M}
P\paren{D \mid M} \propto P\paren{D_{\Pi_x} \mid M} \cdot \prod_{i = 1}^{n_c} P\paren{D_{C_i} \mid M, D_{\boldsymbol{S_i}}}\text{.}
\end{equation}

@@

\noindent
where $n^{(j)}_{i,m,l}$ is the number of instances in $i$th interval with $m$th value of $C_j$ and $l$th value of $\boldsymbol{S_j}$, $n_{i,l} = \sum_{m=1}^{J_j} n^{(j)}_{i,m,l}$, and $n_i = \sum_{l=1}^{L_j} n_{i,l}$.
The two factors on the right hand side come from the third prior: all distribution of values of $C_j$ in a given interval and with a given value of $S_j$ are equiprobable. According to the forth prior, these distributions are independent from each other, and one thus takes their product. If $S_j = \emptyset$, then Equation~\ref{eq:likelihood_two} is equivalent to

@@

The objective function can be formulated given equations \ref{eq:p_M}, \ref{eq:p_D_given_M}, \ref{eq:likelihood_one}, and \ref{eq:likelihood_three}.
The log-inverse of $P(M) \cdot P(D|M)$ is minimized for computational convenience:

@@

Now we are able to solve the optimization problem over Equation~\ref{eq:opt_prob}.
The dynamic programming procedure is shown in Algorithm~\ref{alg:disc_one}.
It takes two inputs: $D$, the joint samples over all variables sorted in ascending order according to $D_X$, and $G$, the network structure.
The runtime of Algorithm \ref{alg:disc_one} is also $O(n_c  {v'_\text{max}} \cdot N^2 + {v'_\text{max}}^{n_p} \cdot N^2)$, because the runtime of the dynamic programming procedure is less than the runtime for computing $h(u,v)$.
Algorithm \ref{alg:disc_one} is guaranteed to be optimal.
For faster methods with suboptimal results please refer to \citep{Boulle_2006}.

@@

Algorithm \ref{alg:disc_one} has an exponential $\bigo\paren{\paren{v'_\text{max}}^{n_p} \cdot N^2}$ runtime which severely limits discretization in networks with a large number of discretization intervals or parent variables.
This section introduces an approximation to the objective function that significantly reduces the runtime and still preserves the quality of discretization.
The approximation replaces the dominator of the last factor in Equation~\ref{eq:likelihood_one} with

@@

The approximation reduces the runtime of Algorithm~\ref{alg:disc_one} to a polynomial, $\bigo\paren{ {v'_\text{max}} \cdot (n_c + n_p) \cdot N^2}$.
It will be shown in Section~\ref{sec:experiments} that the approximation is more sensitive to the distribution over other variables and is biased towards slightly higher interval counts.
\tim{We need to have a more specific thing to say along the lines of When does it fail?}

With the approximated objective function, Equation~\ref{eq:opt_prob_approx}, provides intuition into how child and parent variables contribute to the objective function.
For example, in the left graph of Figure~\ref{fig:example_networks}, the corresponding square brackets in the approximated objective function are

@@

\begin{algorithm}
  \label{alg:structure_learn}
  \caption{ Learning a discrete-valued Bayesian network}
  \begin{algorithmic}[5]
  \Function{Learn\_DBN}{$D''$, $C$, $order$, $u_{parent}$, $C$, $u_{cycle}$}
    \State $n \leftarrow$ the number of variables in the Bayesian network
    \State $k \leftarrow median\{ \|v\|, v\notin C\}$
    \State $M[i] \leftarrow$ the discretization policy for the $i$th variable
    \State $D^*_i \leftarrow M[i] (D_i)$, list of samples $D_i$ discretized according to $M[i]$
    \State $D^* \leftarrow $ the discretized data for all variables, where $D^*_i \leftarrow D''_i$ if $i \notin C$
    \State $G \leftarrow$ the initial edgeless graph structure
    \For {$i = 1$ to $n$}
      \If {$i \in C$}
        \State $M[i] \leftarrow$  equal-width discretization with $k$ intervals
        \State $D^*_i \leftarrow  M[i] (D''_i)$
      \EndIf
    \EndFor
    \For {$i =1$ to $n$}
      \State $P_{old} \leftarrow f(X^*_i,\Pi_{X^*_i})$: Equation~\ref{eq:f_x_p}
      \State OKToProceed $\leftarrow$ \textbf{true}
      \While {OKToProceed \textbf{and} $\|\Pi_{X^*_i}\| < u_{parent}$}
        \State $Y \leftarrow$ an element from the set $order[1:i] \backslash \Pi_X$
        \State $P_{new} \leftarrow f(X^*_i,\Pi_{X^*_i} \cup Y)$
        \If {$P_{new} > P_{old}$}
          \State $P_{old} \leftarrow P_{new}$
          \State $\Pi_{X_i} \leftarrow \Pi_{X_i} \cup Y $
          \State $M \leftarrow$ {DISCRETIZE}({$D''$, $G$, $C$, $u_{cycle}$}): Algorithm \ref{alg:disc_two}
          \State $D* = M(D'')$
        \Else
          \State OKToProceed $\leftarrow$ \textbf{false}
        \EndIf
      \EndWhile
    \EndFor
    \State \Return $G$, $M$
  \EndFunction
  \end{algorithmic}
\end{algorithm}
